---
title: "Animated NFL Play"
output: html_notebook
---

This is an example of how to create an animated data visualization using `plotly` and packages in `tidyverse` (primarily `dplyr` and `ggplot2`) with NFL Next Gen Stats, which translate a player's spatiotemporal information during NFL games onto a coordinate grid.

### Version

Before you get started, here's info on the software and hardware I used to generate this notebook, along with package versions:

```{r}
# system
version
# plotly
packageVersion("plotly")
# tidyverse
packageVersion("tidyverse")
```

The data used here is a single play from the [Kaggle NFL Punt Analytics Competition](https://www.kaggle.com/c/NFL-Punt-Analytics-Competition/data). 

```{r}
input <- readr::read_csv("https://www.github.com/jimtheflash/")
```

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

